
---
title: "Spatial Cluster analysis - Group3"
author: "Gizem Gulsiye Güleli & Ahmed Abdelmaksoud"
date: "2023-11-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,include=FALSE,warning=FALSE, results=FALSE, message=FALSE}
##Loading and reading packages if its required
if(!require(sf)){install.packages("sf")}
if(!require(sp)){install.packages("sp")}
if(!require(spdep)){install.packages("spdep")}
if(!require(RColorBrewer)){install.packages("RColorBrewer")}
if(!require(ggplot2)){install.packages("ggplot2")}
if(!require(satellite)){install.packages("satellite")}
if(!require(CARBayes)){install.packages("CARBayes")}
if(!require(mclust)){install.packages("mclust")}
if(!require(leaflet)){install.packages("leaflet")}
if(!require(leaflet.extras)){install.packages("leaflet.extras")}

```


```{r, echo=T,warning=FALSE,results=FALSE, message=FALSE}
library(sf) #to read shape files and convert to sf files
library(sp) #to convert sp object
library(spdep) #to create neighborhood matrıx &  For spatial dependencies
library(RColorBrewer) #for mapping  
library(ggplot2) #for mapping  
library(satellite) #it was neede to download CARBayes package
library(CARBayes)  # For Bayesian model
library(mclust) #for the bivariate mixture model
library(leaflet)
library(leaflet.extras)

```



#Q1.	Create the combined shape and data file, calculate the Standard Mortality Ratio (SMR) for each year separately, calculate summary statistics and map the risks for the two years in separate maps. Comment on the summaries and any spatial patterns of interest you see in the maps. Check for spatial autocorrelation in each of the years. 

##The data 

There are 2 folders :

•	The Data folder contains 2 csv files containing expected counts and observed counts in the 2 years.
    -expected_counts.csv file have 271 observations with 3 variables that are; code, E2004, E2005
    -respiratory_admissions.csv file have again 271 observations with 3 variables that are; IG, Y2004, Y2005


•	The SG_IntermediateZoneBdry_2001 folder contains the shapefiles for constructing a spatial object in R. 
    -Shape files have 1235 observations and 6 variables which are ; IZ_CODE, IZ_NAME, STDAREA_HA, Shape_Leng, Shape_Area, geometry
 


```{r  read two data set which are called "expected_counts" and "respiratory_admissions"}
data1 <- read.csv(file="Data/expected_counts.csv")
head(data1)

data2 <- read.csv(file="Data/respiratory_admissions.csv")
head(data2)

```

 looks like code and IG variables are matchıng and we can merge the two dataset but first ı wanted to check is there any non-matching value or not to don't miss any value or decide merging method.

```{r}
# Check non-matching values between "code" and "IG" variables
non_matching_values <- setdiff(data1$code, data2$IG)

# Print the non-matching values
if(length(non_matching_values) > 0) {
  print(non_matching_values)
} else {
  print("No non-matching values found.")
}
```

There is no non-matching value thus we can use inner join method to merge these two dataset easily.

```{r}
data <- merge(data1, data2, by.x = "code", by.y = "IG", all = TRUE)
head(data)


```



```{r we read the shapefiles by using "sf" package and printed first 3 rows. }
shape<-read_sf("SG_IntermediateZoneBdry_2001/")
head(shape, 3)
```


We observed that the unique IZ codes appear in this data table too, so that is how we now link the shapefiles to our data using the **merge()** function. But this time we are sure that shapefiles have much more observation then dataset. Thats why we will try to protect dataset observations.

```{r Perform left join to merge shapefiles with dataset}

merged_data <- merge(data, shape, by.x = "code", by.y = "IZ_CODE", all.x = TRUE)
```

The set of data points in the data set can relate to a subset or all of the areas in the shapefile. In this example the shapefile elements contain the polygon information for all IZs in Scotland, where as the data set only contains data on the IZs in  Glasgow. Thus the combined spatial data object **sp.dat** only contains the set of IZs in Glasgow. To observe this we plot the both shape and merged data geometry.

```{r, fig.align='center' }
# Set up a 1x2 plotting layout
par(mfrow = c(1, 2))

# Plot the geometry of the shapefile
plot(shape$geometry, main = "Shapefile Geometry")

# Plot the geometry of the merged dataset
plot(merged_data$geometry, main = "Merged Data Geometry")
```
```{r}
str(merged_data)
class(merged_data)
```
## Calculating SMR and Summary Statistics (including EDA for SMR values )

```{r calculate the Standard Mortality Ratio (SMR) for each year separately, }
# Calculate SMR for 2004
merged_data$SMR_2004 <- (merged_data$Y2004 / merged_data$E2004)

# Calculate SMR for 2005
merged_data$SMR_2005 <- (merged_data$Y2005 / merged_data$E2005)


```


```{r calculate summary statistics }
summary(merged_data)

summary(merged_data[c("SMR_2004", "SMR_2005")])

#EDa for SMR variables
par(mfrow = c(1, 2))

# Plot histogram for SMR_2004 and SMR_2005 to see distribution 

hist(merged_data$SMR_2004, main = "SMR 2004 Distribution", xlab = "SMR_2004", col = "lightblue")
hist(merged_data$SMR_2005, main = "SMR 2005 Distribution", xlab = "SMR_2005")

# box plot for distrubiton and observe outliers
boxplot(merged_data$SMR_2004, main = "SMR 2004 Boxplot", ylab = "SMR_2004", col = "lightblue", border = "black",notch = TRUE )
boxplot(merged_data$SMR_2005, main = "SMR 2005 Boxplot", ylab = "SMR_2005",notch = TRUE)


par(mfrow = c(1, 1))

#scatter plot to observe relation between two variable 
plot(merged_data$SMR_2004, merged_data$SMR_2005, main = "SMR 2004 vs. SMR 2005", xlab = "SMR_2004", ylab = "SMR_2005")
abline(lm(merged_data$SMR_2005 ~ merged_data$SMR_2004), col = "red")

cor(merged_data[c("SMR_2004", "SMR_2005")])


```

 The minimum and maximum values  suggest that the overall range of standardized mortality ratios did not change significantly between the two years (minimum value equals 0.18 and maximum value equals 1.83 for 2004 ,and minimum value equals 0.31 and maximum value equals 1.98 for 2005)

The median and mean values indicate the central tendency of the data, with the mean slightly higher than the median, suggesting a right-skewed distribution for both year. Histogram plot also showing and proving the right-skewed distribution for both year.
Also the values are close to each other for both years, however "SMR_2005" having a slightly higher compared to "SMR_2004"

For both SMR_2004 and SMR_2005, the box plots show the existing of outliers—individual data points beyond the whiskers. These extreme values suggest potential variability in specific areas, prompting further investigation into the factors influencing high standardized mortality ratios.

we also check  the scatter plot to displays the relationship between "SMR_2004" and "SMR_2005" values for each observation.This type of plot is useful for visually inspecting the correlation or pattern between two variables. And from the graph we are able to observe positive relation between two variable.   So we calculated the correlation coefficient between "SMR_2004" and "SMR_2005" which is equal  0.85144 and indicates that there is highly positive linear relationship between the two variables.
The high positive correlation indicates that areas with higher standardized mortality ratios in one year ("SMR_2004") tend to also have higher standardized mortality ratios in the other year ("SMR_2005").
We could suggest some level of consistency or similarity in the factors influencing standardized mortality ratios between the two years.
However correlation result does not imply causation; additional analysis is essential to uncover the underlying factors behind the observed correlation

## Convert Data Types and Create Spatial Objects for further study

```{r  }

# Simplify the multipolygon to a single polygon
merged_data$geometry <- st_simplify(merged_data$geometry)

# Convert to sf object
sf.dat <- st_as_sf(merged_data)

# Convert to sp object
sp.dat <- as(sf.dat, "Spatial")

```


## Mapping the risks for the two years in separate maps

```{r }

# Set up a 1x2 layout for two plots
par(mfrow = c(1, 2))

# for 2004
ggplot(data = sf.dat) +
  geom_sf(aes(fill = SMR_2004)) +
  coord_sf() +
  xlab("Easting (m)") +
  ylab("Northing (m)") +
  labs(title = "Risk for 2004", fill = "SMR") +
  theme(title = element_text(size=16)) +
scale_fill_gradientn(colors=brewer.pal(n=9, name="YlOrRd"))


# for 2005
ggplot(data = sf.dat) +
  geom_sf(aes(fill = SMR_2005)) +
  coord_sf() +
  xlab("Easting (m)") +
  ylab("Northing (m)") +
  labs(title = "Risk for 2005", fill = "SMR") +
  theme(title = element_text(size=16)) +
scale_fill_gradientn(colors=brewer.pal(n=9, name="YlOrRd"))

```


##Comment on the summaries and any spatial patterns of interest you see in the maps. 


SMR > 1: Indicates a higher occurrence of respiratory diseases than expected. This might suggest potential health risks or factors contributing to an elevated incidence of respiratory diseases in the studied population.

SMR < 1: Indicates a lower occurrence of respiratory diseases than expected. This might suggest protective factors or a population that experiences fewer respiratory diseases than what would be anticipated based on the standard population.

SMR = 1: Indicates a respiratory disease occurrence similar to the expected rate. In this case, the observed respiratory disease occurrences align with what would be expected based on the standard population.


We can observe that areas with high SMR values tend to be close to other areas with high SMR values, and areas with low SMR values tend to be close to other areas with low SMR values 




```{r Check for spatial autocorrelation in each of the years.}

# Create neighborhood matrix
W.nb <- poly2nb(sf.dat$geometry, row.names = rownames(sp.dat))
W <- nb2mat(W.nb, style = "B")
W.list <- nb2listw(W.nb, style = "B")

# Moran's I for SMR in 2004
moran_I_2004 <- moran.test(sf.dat$SMR_2004, W.list)
print("Moran's I for SMR in 2004:")
print(moran_I_2004)

# Moran's I for SMR in 2005
moran_I_2005 <- moran.test(sf.dat$SMR_2005, W.list)
print("Moran's I for SMR in 2005:")
print(moran_I_2005)
```
The low p-values and positive autocorrelation suggest that there is a significant spatial clustering of SMR values in both 2004 and 2005.

The positive Moran's I statistic indicates a positive spatial autocorrelation, suggesting that spatial clusters where neighboring regions exhibit similar patterns in respiratory disease occurrence, as indicated by comparable SMR values. This positive spatial autocorrelation observed in the Moran's I statistic is consistent with what is visually evident on the maps. 

The positive autocorrelation suggests the presence of spatial clusters, meaning that neighboring regions have similar SMR values. This could indicate the presence of spatial patterns in health outcomes.



#Q2.	Run a separate Poisson Leroux CAR model on each year for the respiratory counts using only an intercept term (no covariates) and expected count offset. Produce some summaries examining convergence of the chain and comment on them. Plot the fitted values on two maps and comment on them.

##A simple (no correlation) covariate model

```{r  -for 2004}
# Setting the regression formula/equation
formula1 <- Y2004 ~ offset(log(E2004))

# Fit using MLE/glm
model1 <- glm(formula = formula1, family = "poisson", data = sp.dat)

# Summary of the model
summary(model1)

# Check the relative risks
exp(cbind(model1$coefficients, confint(model1)))

# Residual diagnostics - normality and constant variance
par(mfrow = c(1, 2))
qqnorm(residuals(model1, type = "pearson"), main = "", pch = 19)
qqline(residuals(model1, type = "pearson"), col = "red")
plot(model1$fitted.values, residuals(model1, type = "pearson"), pch = 19, 
     xlab = "Fitted values", ylab = "Residuals")

# Residual diagnostics - correlation
# Replace W.list with our actual spatial weights matrix.
moran.mc(x = residuals(model1, type = "pearson"), listw = W.list, nsim = 10000)

```

```{r -for 2005}
# Setting the regression formula/equation for model2
formula2 <- Y2005 ~ offset(log(E2005))

# Fit using MLE/glm for model2
model2 <- glm(formula = formula2, family = "poisson", data = merged_data)

# Summary of model2
summary(model2)

# Check the relative risks for model2
exp(cbind(model2$coefficients, confint(model2)))

# Residual diagnostics - normality and constant variance for model2
par(mfrow = c(1, 2))
qqnorm(residuals(model2, type = "pearson"), main = "", pch = 19)
qqline(residuals(model2, type = "pearson"), col = "red")
plot(model2$fitted.values, residuals(model2, type = "pearson"), pch = 19, 
     xlab = "Fitted values", ylab = "Residuals")

# Residual diagnostics - correlation for model2
# Replace W.list with our actual spatial weights matrix.
moran.mc(x = residuals(model2, type = "pearson"), listw = W.list, nsim = 10000)

```
The results from fitting the Poisson log-linear models for both years (2004 and 2005) show similar patterns.
The intercepts are approximately equals to -0.16 (model1) and -0.19 (model2) and they represents the log of the expected count when all other covariates are zero.

The relative risks for the intercepts are approximately 0.82 (model1)  and 0.86 (model2).

QQ plots and residuals vs. fitted values plots are indicate reasonably normal and constant variance of residuals.

The Moran’s I statistics are approximately equal to 0.40(model1) and 0.43 (model2)  and the p-values are less than 0.05 and  suggests that the significant positive spatial correlation in residuals. Thus the independence assumption made by the models  not adequate. Therefore we now fit a spatial correlation model using the CARBayes package.

##Spatial modelling with CARBayes

```{r - for 2004}

# Fitting the spatial correlation model for 2004
model_2004 <- S.CARleroux(formula = formula1, family = "poisson", data = sf.dat, W = W,
                          burnin = 10000, n.sample = 100000, thin = 10, verbose = FALSE)

# Print model and summary of the model for 2004
print(model_2004)
summary(model_2004)


# Check convergence for 2004
summary(model_2004$samples)
plot(model_2004$samples$rho)

# Assess goodness of fit for 2004
moran.mc(x = residuals(model_2004, type = "pearson"), listw = W.list, nsim = 10000)


```


```{r}

# Inference from the model for 2004

# Estimated relative risks and 95% credible intervals
exp(model_2004$summary.results[, c("Mean", "2.5%", "97.5%")])

# Calculate  risk for 2004
sp.dat@data$risk_2004 <- model_2004$fitted.values / sp.dat@data$E2004


# Calculate Posterior Exceedance Probability (PEP) for 2004
m <- nrow(model_2004$samples$fitted)
risk_2004 <- model_2004$samples$fitted / matrix(rep(sp.dat@data$E2004, nrow(model_2004$samples$fitted)), nrow = m, byrow = TRUE)
summarise.exceedences<-function(samples,exceedences)
{
exceed.vec<-apply((samples>exceedences),2,sum)/nrow(samples)
9
return(exceed.vec)
}

sp.dat@data$PEP_2004 <- as.numeric(summarise.exceedences(risk_2004, 1))


##mapping the results for 2004 
sp.dat.ll <- spTransform(sp.dat, CRS("+proj=longlat +datum=WGS84 +no_defs"))

par(mfrow = c(1, 2))

##Risk 

colours_risk_2004 <- colorNumeric(palette = "YlOrRd", domain = sp.dat.ll@data$risk)


leaflet(data = sp.dat.ll) %>%
  addTiles() %>%
  addPolygons(fillColor = ~colours_risk_2004(risk_2004),
              color = "black", weight = 1,
              fillOpacity = 0.7) %>%
  addLegend(pal = colours_risk_2004, values = sp.dat.ll@data$risk_2004,
            opacity = 1, title = "Risk 2004") %>%
  addScaleBar(position = "bottomleft")

##PEP
colours_pep_2004 <- colorNumeric(palette = "YlOrRd", domain = sp.dat.ll@data$PEP_2004)

leaflet(data=sp.dat.ll) %>%
  addTiles() %>%
  addPolygons(fillColor = ~colours_pep_2004(PEP_2004),
              color="black", weight=1,
              fillOpacity = 0.7) %>%
  addLegend(pal = colours_pep_2004, values = sp.dat.ll@data$PEP_2004,
            opacity = 1, title="PEP 2004 ") %>%
  addScaleBar(position="bottomleft")

```





```{r Spatial modelling with CARBayes- for 2005}

# Fitting the spatial correlation model for 2005
model_2005 <- S.CARleroux(formula = formula2, family = "poisson", data = sf.dat, W = W,
                          burnin = 10000, n.sample = 100000, thin = 10, verbose = FALSE)

# Print model and summary of the model for 2005
print(model_2005)
summary(model_2005)

# Check convergence for 2005
summary(model_2005$samples)
plot(model_2005$samples$rho)

# Assess goodness of fit for 2005
moran.mc(x = residuals(model_2005, type = "pearson"), listw = W.list, nsim = 10000)


```


```{r}
# Inference from the model for 2005
# Estimated relative risks and 95% credible intervals
exp(model_2005$summary.results[, c("Mean", "2.5%", "97.5%")])

# Calculate risk for 2005
sp.dat@data$risk_2005 <- model_2005$fitted.values / sp.dat@data$E2005

# Calculate Posterior Exceedance Probability (PEP) for 2005
m <- nrow(model_2005$samples$fitted)
risk_2005 <- model_2005$samples$fitted / matrix(rep(sp.dat@data$E2005, nrow(model_2005$samples$fitted)), nrow = m, byrow = TRUE)
sp.dat@data$PEP_2005 <- as.numeric(summarise.exceedences(risk_2005, 1))

## Mapping the results for 2005
sp.dat.ll <- spTransform(sp.dat, CRS("+proj=longlat +datum=WGS84 +no_defs"))

par(mfrow = c(1, 2))

## Risk
colours_risk_2005 <- colorNumeric(palette = "YlOrRd", domain = sp.dat.ll@data$risk_2005)

leaflet(data = sp.dat.ll) %>%
  addTiles() %>%
  addPolygons(fillColor = ~colours_risk_2005(risk_2005),
              color = "black", weight = 1,
              fillOpacity = 0.7) %>%
  addLegend(pal = colours_risk_2005, values = sp.dat.ll@data$risk_2005,
            opacity = 1, title = "Risk 2005") %>%
  addScaleBar(position = "bottomleft")

## PEP
colours_pep_2005 <- colorNumeric(palette = "YlOrRd", domain = sp.dat.ll@data$PEP_2005)

leaflet(data = sp.dat.ll) %>%
  addTiles() %>%
  addPolygons(fillColor = ~colours_pep_2005(PEP_2005),
              color = "black", weight = 1,
              fillOpacity = 0.7) %>%
  addLegend(pal = colours_pep_2005, values = sp.dat.ll@data$PEP_2005,
            opacity = 1, title = "PEP 2005") %>%
  addScaleBar(position = "bottomleft")

```
In examining the spatial epidemiological data for 2004 and 2005, Poisson regression models with Leroux CAR spatial correlation were employed, considering offsets of log(E2004) and log(E2005), respectively.
For the 2004 model, the estimated parameters, including the intercept (-0.2624), tau2 (0.3072), and rho (0.7407), indicated a significant spatial correlation. Convergence diagnostics and Moran's I test affirmed model reliability, with spatial risk maps illustrating the distribution of estimated risks across the area.
In contrast, the 2005 model demonstrated a weaker spatial correlation (rho: 0.6764), with parameters - intercept (-0.2114) and tau2 (0.2953) - suggesting increased overall risk. Similar convergence diagnostics and Moran's I tests were applied, emphasizing changes in risk distribution through spatial risk maps.
Comparing the two years, differences in estimated relative risks and spatial correlation emerged. Risk maps for each year aided in visually assessing changing risk distributions. Furthermore, areas with potentially elevated risks, identified through Posterior Exceedance Probability (PEP) maps, offered additional insights into hotspot locations.
Notably, the 2004 model exhibited a lower Deviance Information Criterion (DIC), indicating a better fit compared to the 2005 model. Overall, this spatial epidemiological analysis provided insights into risk dynamics, highlighting potential shifts and hotspots across the study area over the two years.

#Q3.	Extract the mean fitted values (after burn-in) for the areas’ risks in each year. Run a bivariate mixture model clustering method on the fitted values for both years (i.e. cluster both years’ fitted values at the same time, don’t run a separate clustering on each year). Comment on your result in terms of number of clusters chosen, the component means. (Extra credit: map the cluster membership (i.e. one colour for each cluster)).


```{r  }
# Extract mean fitted values for 2004 and 2005
mean_fitted_values_2004 <- colMeans(model_2004$samples$fitted)
mean_fitted_values_2005 <- colMeans(model_2005$samples$fitted)

# Combine the mean fitted values for both years
mean_fitted_values <- cbind(mean_fitted_values_2004, mean_fitted_values_2005)

# Run a bivariate mixture model clustering method
bivariate_model <- Mclust(mean_fitted_values, G = 1:10)  # possible to  adjust the range for the number of clusters (G)

# Summary of the clustering result
summary(bivariate_model)


# Plot the clustering result
plot(bivariate_model, what = "classification")


```


```{r}

library(clustMD)
# Fit a mixture model with different covariance structures
clust_eee <- Mclust(mean_fitted_values, G = 1:10, modelNames = "EEE")
clust_eev <- Mclust(mean_fitted_values, G = 1:10, modelNames = "EII")
clust_vev <- Mclust(mean_fitted_values, G = 1:10, modelNames = "VII")
clust_eei <- Mclust(mean_fitted_values, G = 1:10, modelNames = "EEI")
clust_vei <- Mclust(mean_fitted_values, G = 1:10, modelNames = "VEI")
clust_evi <- Mclust(mean_fitted_values, G = 1:10, modelNames = "EVI")
clust_vvi <- Mclust(mean_fitted_values, G = 1:10, modelNames = "VVI")

# Summary of the clustering results using Mclust
summary(clust_eee)
summary(clust_eev)
summary(clust_vev)
summary(clust_eei)
summary(clust_vei)
summary(clust_evi)
summary(clust_vvi)

# Plot 
plot(clust_eev, what = "classification", main = "EII Covariance")
plot(clust_vev, what = "classification", main = "VII Covariance")
plot(clust_eee, what = "classification", main = "EEE Covariance")
plot(clust_eei, what = "classification", main = "EEI Covariance")
plot(clust_vei, what = "classification", main = "VEI Covariance")
plot(clust_evi, what = "classification", main = "EVI Covariance")
plot(clust_vvi, what = "classification", main = "VVI Covariance")

```


```{r}
# Extract BIC values
bic_values <- c(
  BIC(clust_eee),
  BIC(clust_eev),
  BIC(clust_vev),
  BIC(clust_eei),
  BIC(clust_vei),
  BIC(clust_evi),
  BIC(clust_vvi)
)

# Display BIC values
cat("BIC Values:\n")
cat("EEE Covariance:", BIC(clust_eee), "\n")
cat("EII Covariance:", BIC(clust_eev), "\n")
cat("VII Covariance:", BIC(clust_vev), "\n")
cat("EEI Covariance:", BIC(clust_eei), "\n")
cat("VEI Covariance:", BIC(clust_vei), "\n")
cat("EVI Covariance:", BIC(clust_evi), "\n")
cat("VVI Covariance:", BIC(clust_vvi), "\n")

# Choose the best model
best_model <- switch(
  which.min(bic_values),
  "EEE Covariance" = clust_eee,
  "EII Covariance" = clust_eev,
  "VII Covariance" = clust_vev,
  "EEI Covariance" = clust_eei,
  "VEI Covariance" = clust_vei,
  "EVI Covariance" = clust_evi,
  "VVI Covariance" = clust_vvi
)

# Plot the best model
plot(best_model, what = "classification", main = "Best Model")

```
```{r compare the best model from clusMD and mclust}
par(mfrow = c(1, 2))

# Plot the best model
plot(best_model, what = "classification")
title("Best Model")

# Plot the original bivariate model
plot(bivariate_model, what = "classification")
title("Original Bivariate Model")

```


## Map the Cluster Membership
```{r}

# Map the cluster membership
cluster_membership <- predict(bivariate_model)$classification

# Create a new data frame with cluster membership and spatial information
cluster_data <- data.frame(ID = rownames(sp.dat@data), Cluster = cluster_membership)

# Merge using base R merge function
sp.dat_cluster <- merge(x = sp.dat.ll, y = cluster_data, by.x = "code", by.y = "ID", all.x = TRUE)

# Mapping the cluster membership
colours_cluster <- colorFactor(palette = "Set1", levels = unique(cluster_membership))

leaflet(data = sp.dat_cluster) %>%
  addTiles() %>%
  addPolygons(fillColor = ~colours_cluster(Cluster),
              color = "black", weight = 1,
              fillOpacity = 0.7) %>%
  addLegend(pal = colours_cluster, values = sp.dat_cluster$Cluster,
            opacity = 1, title = "Cluster Membership") %>%
  addScaleBar(position = "bottomleft")
```


```{r  Bivariate Mixture Model Clustering}

# Fit the Gaussian Mixture Model on mean fitted values
gmm_model <- Mclust(mean_fitted_values, G = 1:5)  

# Get the cluster assignments
cluster_assignments <- gmm_model$classification

# Comment on the results
cat("Number of clusters chosen:", gmm_model$G, "\n")

# Component means
cat("Component means:")
print(gmm_model$means)


```



```{r}

# Map the cluster membership
sf.dat$Cluster <- as.factor(cluster_assignments)

# Define colors for clusters
cluster_colors <- colorFactor(palette = "Set3", levels = unique(cluster_assignments))

# Create a leaflet map

leaflet(data = sp.dat) %>% 
  addTiles() %>% 
  addPolygons(fillColor = ~cluster_colors(Cluster), 
              color = "white", weight = 1, 
              fillOpacity = 0.7) %>%
  addLegend(pal = cluster_colors, values = sp.dat$Cluster, 
            opacity = 1, title = "Cluster") %>%
  addScaleBar(position = "bottomleft")

```


#Q4:4.	Run another clustering algorithm on the data and use an appropriate measure to compare the two clustering results. (The other algorithm could be k-means, hierarchical clustering, another type of mixture model clustering. The data used could be the raw SMRs or the average fitted values of the risks for the areas for both years, whatever you prefer. You do not need to comment on the other clustering results beyond comparing to the first clustering results from Question 3.)




